---
title: "NHL Player Analysis"
output: 
  html_document:
    theme: united
    highlight: tango
---

# Load necessary libraries
```{r echo=TRUE, message=FALSE}
library(dplyr)
library(ggplot2)
library(tidyr)
```

# Read the data
```{r}
canadian_players <- read.csv("canadian_players.csv")
all_players <- read.csv("all_players.csv")
```

## Explore and Visualize the Distribution of Birth Quarters

### Distribution for All Players
```{r echo=TRUE}
ggplot(all_players, aes(x = factor(BirthQuarter))) +
  geom_bar(fill = "skyblue", color = "black") +
  labs(title = "Distribution of Birth Quarters for All Players", y = "Count") +
  theme_minimal()
```

### Distribution for Canadian Players
```{r echo=TRUE}
ggplot(canadian_players, aes(x = factor(BirthQuarter))) +
  geom_bar(fill = "lightgreen", color = "black") +
  labs(title = "Distribution of Birth Quarters for Canadian Players", y = "Count") +
  theme_minimal()
```

The bar plots do not show a significant difference visually.

## Hypothesis Test for Birth Quarters

### Percentage of Each Birth Quarter
```{r echo=TRUE}
percent_table_all <- table(all_players$BirthQuarter) / nrow(all_players) * 100
percent_table_canadian <- table(canadian_players$BirthQuarter) / nrow(canadian_players) * 100

two_way_table <- cbind(percent_table_all, percent_table_canadian)
colnames(two_way_table) <- c("All Players", "Canadian Players")
two_way_table
```

### Hypothesis Test (Chi-squared)
```{r echo=TRUE}
test_result <- chisq.test(two_way_table)
test_result
```

Fail to reject the null hypothesis; the birth quarter distributions are not statistically different.

## Explore Player Positions

### Relative Tables
```{r echo=TRUE}
relative_table_all <- table(all_players$Position) / nrow(all_players) * 100
relative_table_canadian <- table(canadian_players$Position) / nrow(canadian_players) * 100

relative_table_all
relative_table_canadian
```

No statistically significant differences in player positions for Canadian players.

### Two-way Table for Player Positions
```{r echo=TRUE}
position_table_all <- table(all_players$Position)
position_table_canadian <- table(canadian_players$Position)

percent_table_all_positions <- position_table_all / sum(position_table_all) * 100
percent_table_canadian_positions <- position_table_canadian / sum(position_table_canadian) * 100

two_way_table_positions <- cbind(percent_table_all_positions, percent_table_canadian_positions)
colnames(two_way_table_positions) <- c("All Players", "Canadian Players")
two_way_table_positions
```

### Hypothesis Test for Player Positions (Chi-squared)
```{r echo=TRUE}
test_result_positions <- chisq.test(two_way_table_positions)
test_result_positions
```

Fail to reject the null hypothesis; the position distributions are not statistically different.

## Analyze Birth Quarter Frequency

### Table for Birth Quarters in Canadian Players
```{r echo=TRUE}
quarter_table_canadian <- table(canadian_players$BirthQuarter)
quarter_table_canadian
```

### Chi-squared Test for Birth Quarters in Canadian Players
```{r echo=TRUE}
chi_square_result <- chisq.test(quarter_table_canadian)
chi_square_result
```

Reject the null hypothesis; a statistically significant difference in birth quarter distribution for Canadian players.

### Percentage Table for Birth Quarters in Canadian Players
```{r echo=TRUE}
percent_table_canadian <- quarter_table_canadian / sum(quarter_table_canadian) * 100
percent_df_canadian <- data.frame(Quarter = as.character(seq_along(percent_table_canadian)), Percentage = percent_table_canadian)
percent_df_canadian
```

There is a cascade pattern in birth quarter frequency.

---

This analysis indicates a potential impact of birth quarters on the likelihood of players making it to the NHL, especially for Canadian players.